---
title: "Untitled"
output: html_document
fig_width: 8
fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set new parameters for reproducible pipeline

```{r}
# REQUIRED PARAMETERS
course <- 'BIOL_228_4001_4002'
course_directory <- paste0("/Users/bencroft/", course, '/input/')

# OPTIONAL PARAMETERS:

# Option 1: Combine two course sections
#course1_directory <- 
#course2_directory <- 
#course_combined_directory <- 
#course1_section_number <- 
#course2_section_number <- 
```

## Load libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(scales)
library(stringr)
library(ggrepel)
library(ggridges)
```

## Load helper scripts (functions)

```{r}
source("./scripts/wrangling.R")
source("./scripts/visualizer.R")
source("./scripts/section_combiner.R")
```

## Load data

### Set data paths

```{r}
dynamics_path <- paste0(course_directory, "dynamics.csv")
grades_path <-paste0(course_directory, "grades.csv")
items_path <- paste0(course_directory, "items.csv")
statics_path <- paste0(course_directory, "statics.csv")
submissions_path <- paste0(course_directory, "submissions.csv")
sis_path <- paste0(course_directory, "sis.csv")
```

### Read data into memory

```{r}
dynamics_input    <- read.csv(dynamics_path)
grades_input      <- read.csv(grades_path)
items_input       <- read.csv(items_path)
statics_input     <- read.csv(statics_path)
submissions_input <- read.csv(submissions_path)
sis_input         <- read.csv(sis_path)
```

```{r}
dynamics <- dynamics_input
grades <- grades_input
items <- items_input
statics <- statics_input
submissions <- submissions_input
sis_acad_plan <- sis_input
sis_no_acad_plan <- sis_input
```

## Process data

### Remove duplicates from `sis_no_acad_plan` due to plan and subplan column removal

```{r}
# Find duplicates in students - largely due to multiple AcademicPlan values per student
# sis %>%
#   select(-AcademicSubPlan, -AcademicPlan) %>%
#   distinct() %>%
#   group_by(StudentID_Anon) %>%
#   summarise(count = n()) %>%
#   arrange(desc(count)) 

sis_no_acad_plan <- sis_no_acad_plan %>%
  select(-AcademicPlan, -AcademicSubPlan) %>%
  distinct()
```


## Process sis_acad_plan

### Get plans with 5+ students

```{r}
plans_5_or_more <- sis_acad_plan %>%
  group_by(AcademicPlan) %>%
  summarise(c = n()) %>%
  arrange(desc(c)) %>%
  filter(c >= 5)

plans_5_or_more
```

### Filter for top plans

```{r}
sis_acad_plan <- sis_acad_plan %>%
  filter(AcademicPlan %in% plans_5_or_more$AcademicPlan) %>%
  select(-AcademicSubPlan) %>%
  distinct()
```


### Fix time formatting in `dynamics`

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
dynamics <- fix_time_band_24hr(dynamics)
dynamics <- fix_week_level(dynamics)
dynamics <- fix_day_of_week(dynamics)
```

### Add space in BISQL student names

```{r}
sis_acad_plan %>%
  mutate(StudentName = addSpace(StudentName))
sis_no_acad_plan %>%
  mutate(StudentName = addSpace(StudentName))
```

```{r}
## Ensure equal lengths among datasets
# length(unique(dynamics$StudentName))
# length(unique(grades$StudentName))
# length(unique(items$StudentName))
# length(unique(sis_acad_plan$StudentName))
# length(unique(sis_no_acad_plan$StudentName))
# length(unique(statics$StudentName))
# length(unique(submissions$StudentName))
```

### Fix data type in `sis_acad_plan` and `sis_no_acad_plan`

```{r}
sis_acad_plan <- sis_acad_plan %>%
  mutate(StudentID = as.character(StudentID))
sis_no_acad_plan <- sis_no_acad_plan %>%
  mutate(StudentID = as.character(StudentID))
```

### Anonymize student identifiers

#### Create StudentID column

```{r}
dynamics <- create_student_id_col(dynamics)
grades <- create_student_id_col(grades)
items <- create_student_id_col(items)
statics <- create_student_id_col(statics)
```

#### Create anonymous/random ids

```{r}
set.seed(1)

student_roster <- union(dynamics$StudentID, grades$StudentID)       # Intersect dynamics and grades
student_roster <- union(student_roster, items$StudentID)            # Intersect items    
student_roster <- union(student_roster, statics$StudentID)          # Intersect statics
student_roster <- union(student_roster, sis_acad_plan$StudentID)    # Intersect sis_acad_plan
student_roster <- union(student_roster, sis_no_acad_plan$StudentID) # Intersect sis_no_acad_plan


# Create hash table for anonymous student ids
student_count <- length(unique(student_roster))
student_anon_ids <- sample(10000:99999, student_count, replace=FALSE)
student_anon_df <- data.frame(student_roster, student_anon_ids) %>%
  rename(StudentID = student_roster, StudentID_Anon = student_anon_ids)

dynamics <- left_join(dynamics, student_anon_df, by = c("StudentID"))
grades <- left_join(grades, student_anon_df, by = c("StudentID"))
items <- left_join(items, student_anon_df, by = c("StudentID"))
statics <- left_join(statics, student_anon_df, by = c("StudentID"))
sis_acad_plan <- left_join(sis_acad_plan, student_anon_df, by = c("StudentID"))
sis_no_acad_plan <- left_join(sis_no_acad_plan, student_anon_df, by = c("StudentID"))
```

#### Remove PII

```{r}
dynamics <- dynamics %>%
  select(-StudentName, -StudentID)
grades <- grades %>%
  select(-StudentName, -StudentID)
items <- items %>%
  select(-StudentName, -StudentID)
statics <- statics %>%
  select(-StudentName, -StudentID)
sis_acad_plan <- sis_acad_plan %>%
  select(-StudentName, -StudentID, -EmployeeID)
sis_no_acad_plan <- sis_no_acad_plan %>%
  select(-StudentName, -StudentID, -EmployeeID)
```

## Visualization

### Outcomes

#### Final Grade Distribution

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_FinalGradeDistribution(statics)
```

#### Outcomes by Student Credit Load

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByStudentCreditLoad(statics, sis_no_acad_plan)
```

#### Outcomes by Gender

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByGender(statics, sis_no_acad_plan)
```

#### Outcomes by First Time Status

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByFirstTermAtInstitution(statics, sis_no_acad_plan)
```

#### Outcomes by Academic Level

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByAcademicLevel(statics, sis_no_acad_plan)
```

#### Outcomes by First Gen

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByFirstGeneration(statics, sis_no_acad_plan)
```


### Outcomes by Plan

```{r}
viz_OutcomesByFirstGeneration <- function(statics, sis_acad_plan) {
  
  # Create plot
  tmp <- statics %>%
    left_join(sis_no_acad_plan, by = "StudentID_Anon") %>%
    filter(!SISGradeLetter %in% c('W', 'CW', 'I')) %>%
    mutate(FirstGen = case_when(FirstGen == "First Generation (Y)" ~ "First Gen",
                                FirstGen == "Not First Generation (N)" ~ "Not First Gen",
                                FirstGen == "Unknown" ~ "Unknown"))
  
  g <- ggplot(tmp, aes(y = NormalizedScore, x = as.factor(FirstGen))) +
    geom_boxplot(aes(fill = as.factor(FirstGen)), outlier.shape = NA) +
    geom_jitter(width = 0.10) +
    my_theme() +
    labs(title = "Course Outcomes by First Generation Status", 
         y = "Course Grade (Percent)", 
         x = "", 
         color = "Credit Load",
         subtitle = "Summer 2020 BIOL 228 | Section 4001 and 4002",
         caption = "Note: This dataset excludes students who withdrew or had an incomplete term") +
    scale_fill_manual(values = c("#009DD9", "#999999", "#999999", "#999999", "#999999")) +
    theme(legend.position = "") +
    #scale_x_discrete(labels = abbreviate) +
    scale_y_continuous(limits = c(0, 100))
  
}

v <- viz_OutcomesByAcademicPlan(statics, sis_acad_plan)
```






### Item Grades

```{r, fig.height = 11, fig.width = 8}
v <- viz_CourseItemGradeDistributions(grades)
v
```

## Bb Activity by Group

### Bb Activity by SISGradeLetter

#### Table: All

```{r}
t <- table_ActivityBySISGradeLetter(statics)
```

#### Boxplot: Hours

```{r, fig.width = 8, fig.height = 7}
v <- viz_Activity_HoursBySISGradeLetter(statics)
```

#### Boxplot: Logins

```{r}
v <- viz_Activity_LoginsBySISGradeLetter(statics)
```

#### Boxplot: Post Characters

```{r}
v <- viz_Activity_CharactersBySISGradeLetter(statics)
```

### Bb Activity by FirstGen

#### Table: All

```{r}
t <- table_ActivityByFirstGen(statics, sis_no_acad_plan)
```

#### Boxplot: Hours

```{r}
v <- viz_Activity_HoursByFirstGen(statics, sis_no_acad_plan)
```

#### Boxplot: Logins

```{r}
v <- viz_Activity_LoginsByFirstGen(statics, sis_no_acad_plan)
```

#### Boxplot: Post Characters

```{r}
v <- viz_Activity_CharactersByFirstGen(statics, sis_no_acad_plan)
```


# ********************

## Prerequisite analysis

### Load data

#### Set data paths

```{r}
prerequisites_path <- paste0(course_directory, "prerequisites.csv")
```

#### Read data into memory

```{r}
prerequisites_input    <- read.csv(prerequisites_path)
```

```{r}
prerequisites <- prerequisites_input
```

```{r}
colnames(prerequisites) <- c("TermKey", "TermSourceKey", "SessionCode", "ClassNumberSectionUniqueDescription", "ClassNumberUniqueDescription",
                             "PrimarySubject", "PrimaryCatalogNumber", "ClassSection", "PrimaryComponent", "CourseAttribute", "CourseAttributeValue",
                             "ValueFormalDescription", "StudentID", "EnrolledClassCount", "DropCount", "WithdrawCount", "CreditsAttempted",
                             "CreditsEarned", "EarnCreditIndicator", "SuccessInidicator", "HassClassGrade", "ClassGrade", "GradeKey",
                             "GradeLetter", "GradeDescription", "GradeSubgroup", "GradeGroup", "GradingBasisDescription")
```

```{r}
head(prerequisites)
```

```{r}
# Need to get grade percent


```



































































