---
title: "Untitled"
output: html_document
fig_width: 8
fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set new parameters for reproducible pipeline

```{r}
# REQUIRED PARAMETERS
course <- '1209_ACCT_205'
course_directory <- paste0("/Users/bencroft/", course, '/input/')

# OPTIONAL PARAMETERS:

# Option 1: Combine two course sections
#course1_directory <- 
#course2_directory <- 
#course_combined_directory <- 
#course1_section_number <- 
#course2_section_number <- 
```

## Load libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(data.table)
library(scales)
library(stringr)
library(ggrepel)
library(ggridges)
library(ggpmisc)
library(kableExtra)
```

## Load helper scripts (functions)

```{r}
source("./scripts/wrangling.R")
source("./scripts/visualizer.R")
source("./scripts/section_combiner.R")
```

## Load data

### Set data paths

```{r}
dynamics_path <- paste0(course_directory, "dynamics.csv")
grades_path <-paste0(course_directory, "grades.csv")
items_path <- paste0(course_directory, "items.csv")
statics_path <- paste0(course_directory, "statics.csv")
submissions_path <- paste0(course_directory, "submissions.csv")
sis_path <- paste0(course_directory, "sis.csv")
```

### Read data into memory

```{r}
dynamics_input    <- read.csv(dynamics_path)
grades_input      <- read.csv(grades_path)
items_input       <- read.csv(items_path)
statics_input     <- read.csv(statics_path)
submissions_input <- read.csv(submissions_path)
sis_input         <- read.csv(sis_path)
```



```{r}
dynamics <- dynamics_input
grades <- grades_input
items <- items_input
statics <- statics_input
submissions <- submissions_input
sis_acad_plan <- sis_input
sis_no_acad_plan <- sis_input
```

## Process data

### A4L Data

### Fix time formatting in `dynamics`

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
dynamics <- fix_time_band_24hr(dynamics)
dynamics <- fix_week_level(dynamics)
dynamics <- fix_day_of_week(dynamics)
```

## Process sis_acad_plan

### Remove duplicates from `sis_no_acad_plan` due to plan and subplan column removal

```{r}
# Find duplicates in students - largely due to multiple AcademicPlan values per student
# sis %>%
#   select(-AcademicSubPlan, -AcademicPlan) %>%
#   distinct() %>%
#   group_by(StudentID_Anon) %>%
#   summarise(count = n()) %>%
#   arrange(desc(count)) 

sis_no_acad_plan <- sis_no_acad_plan %>%
  select(-AcademicPlan, -AcademicSubPlan) %>%
  select(-CumulativeGPA, -HasTransferCumGPA, -RegistrationAddDate, -RegistrationDropDate, -FullTimePartTimeDescription) %>%
  distinct()
```

```{r}
sis_acad_plan <- sis_acad_plan %>%
  select(-CumulativeGPA, -HasTransferCumGPA, -RegistrationAddDate, -RegistrationDropDate, -FullTimePartTimeDescription)
```
















### Get plans with 5+ students

```{r paged.print=TRUE}
plan_counts <- sis_acad_plan %>%
  select(StudentID, AcademicPlan) %>%
  distinct() %>%
  group_by(StudentID, AcademicPlan) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count), AcademicPlan, StudentID) %>%
  ungroup() %>%
  group_by(AcademicPlan) %>%
  summarise(StudentsWithPlan = sum(Count)) %>%
  arrange(desc(StudentsWithPlan))
plan_counts
```

```{r}
plans_5_or_more <- plan_counts %>%
  filter(StudentsWithPlan >= 5)
plans_5_or_more
```

### Filter for top plans

```{r}
sis_acad_plan <- sis_acad_plan %>%
  filter(AcademicPlan %in% plans_5_or_more$AcademicPlan) %>%
  select(-AcademicSubPlan) %>%
  distinct()
```




### Add space in BISQL student names

```{r}
sis_acad_plan %>%
  mutate(StudentName = addSpace(StudentName))
sis_no_acad_plan %>%
  mutate(StudentName = addSpace(StudentName))
```

```{r}
## Ensure equal lengths among datasets
# length(unique(dynamics$StudentName))
# length(unique(grades$StudentName))
# length(unique(items$StudentName))
# length(unique(sis_acad_plan$StudentName))
# length(unique(sis_no_acad_plan$StudentName))
# length(unique(statics$StudentName))
# length(unique(submissions$StudentName))
```

### Fix data type in `sis_acad_plan` and `sis_no_acad_plan`

```{r}
sis_acad_plan <- sis_acad_plan %>%
  mutate(StudentID = as.character(StudentID))
sis_no_acad_plan <- sis_no_acad_plan %>%
  mutate(StudentID = as.character(StudentID))
```

### Anonymize student identifiers

#### Create StudentID column

```{r}
dynamics <- create_student_id_col(dynamics)
grades <- create_student_id_col(grades)
items <- create_student_id_col(items)
statics <- create_student_id_col(statics)
```



#### Create anonymous/random ids

```{r}
set.seed(1)

student_roster <- union(dynamics$StudentID, grades$StudentID)       # Intersect dynamics and grades
student_roster <- union(student_roster, items$StudentID)            # Intersect items    
student_roster <- union(student_roster, statics$StudentID)          # Intersect statics
student_roster <- union(student_roster, sis_acad_plan$StudentID)    # Intersect sis_acad_plan
student_roster <- union(student_roster, sis_no_acad_plan$StudentID) # Intersect sis_no_acad_plan


# Create hash table for anonymous student ids
student_count <- length(unique(student_roster))
student_anon_ids <- sample(10000:99999, student_count, replace=FALSE)
student_anon_df <- data.frame(student_roster, student_anon_ids) %>%
  rename(StudentID = student_roster, StudentID_Anon = student_anon_ids)

dynamics <- left_join(dynamics, student_anon_df, by = c("StudentID"))
grades <- left_join(grades, student_anon_df, by = c("StudentID"))
items <- left_join(items, student_anon_df, by = c("StudentID"))
statics <- left_join(statics, student_anon_df, by = c("StudentID"))
sis_acad_plan <- left_join(sis_acad_plan, student_anon_df, by = c("StudentID"))
sis_no_acad_plan <- left_join(sis_no_acad_plan, student_anon_df, by = c("StudentID"))
```

```{r}
sis_acad_plan
```

```{r}
sym_diff <- function(a,b) setdiff(union(a,b), intersect(a,b))

sis_no_acad_plan$StudentID_Anon %in% student_anon_df$StudentID_Anon
```


#### Remove PII

```{r}
dynamics <- dynamics %>%
  select(-StudentName, -StudentID)
grades <- grades %>%
  select(-StudentName, -StudentID)
items <- items %>%
  select(-StudentName, -StudentID)
statics <- statics %>%
  select(-StudentName, -StudentID)
sis_acad_plan <- sis_acad_plan %>%
  select(-StudentName, -StudentID, -EmployeeID)
sis_no_acad_plan <- sis_no_acad_plan %>%
  select(-StudentName, -StudentID, -EmployeeID)
```

## *******************************





## Visualization

### Outcomes

#### Final Grade Distribution

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_FinalGradeDistribution(statics)
v
```

#### Outcomes by Student Credit Load

```{r}
left_join(statics, sis_no_acad_plan, by = "StudentID_Anon")
```

```{r}
student_anon_df %>%
  filter(StudentID_Anon == 91579)
```



```{r}
sis_input %>%
  filter(StudentID == 114091937)
```


#### Outcomes by Gender

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByGender(statics, sis_no_acad_plan)
v
```

#### Outcomes by First Time Status

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByFirstTermAtInstitution(statics, sis_no_acad_plan)
v
```

#### Outcomes by Academic Level

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByAcademicLevel(statics, sis_no_acad_plan)
v
```

#### Outcomes by First Gen

```{r, fig.retina = 2, fig.width = 8, fig.height = 7}
v <- viz_OutcomesByFirstGeneration(statics, sis_no_acad_plan)
v
```


### Outcomes by Plan

```{r, fig.height = 7, fig.width = 8}
v <- viz_OutcomesByAcademicPlan(statics, sis_acad_plan)
v
```






### Item Grades

```{r, fig.height = 12, fig.width = 9}
v <- viz_CourseItemGradeDistributions(grades)
v
```


# ******

## Bb Activity by Group

### Bb Activity by SISGradeLetter

#### Table: All

```{r}
t <- table_ActivityBySISGradeLetter(statics)
```

#### Boxplot: Hours

```{r, fig.width = 8, fig.height = 7}
v <- viz_Activity_HoursBySISGradeLetter(statics)
v
```

#### Boxplot: Logins

```{r}
v <- viz_Activity_LoginsBySISGradeLetter(statics)
v
```

#### Boxplot: Post Characters

```{r}
v <- viz_Activity_CharactersBySISGradeLetter(statics)
v
```

### Bb Activity by FirstGen

#### Table: All

```{r}
t <- table_ActivityByFirstGen(statics, sis_no_acad_plan)
t %>%
  kable()
```


```{r}
statics %>%
    left_join(sis_no_acad_plan, by = "StudentID_Anon") %>%
    group_by(FirstGen) %>%
    filter(!SISGradeLetter %in% c('W', 'CW', 'I')) %>%
  select(StudentID_Anon) %>%
  table()
```




#### Boxplot: Hours

```{r}
v <- viz_Activity_HoursByFirstGen(statics, sis_no_acad_plan)
v
```

#### Boxplot: Logins

```{r}
v <- viz_Activity_LoginsByFirstGen(statics, sis_no_acad_plan)
v
```

#### Boxplot: Post Characters

```{r}
v <- viz_Activity_CharactersByFirstGen(statics, sis_no_acad_plan)
```


# ********************

## Prerequisite analysis

### Load data

#### Set data paths

```{r}
prerequisites_path <- paste0(course_directory, "prerequisites.csv")
```

#### Read data into memory

```{r}
prerequisites_input <- read.csv(prerequisites_path)
```

```{r}
prerequisites <- prerequisites_input
```

### Process data

#### Fix column names

```{r}
colnames(prerequisites) <- c("TermKey", "TermSourceKey", "SessionCode", "ClassNumberSectionUniqueDescription", "ClassNumberUniqueDescription",
                             "PrimarySubject", "PrimaryCatalogNumber", "ClassSection", "PrimaryComponent", "CourseAttribute", "CourseAttributeValue",
                             "ValueFormalDescription", "StudentID", "EnrolledClassCount", "DropCount", "WithdrawCount", "CreditsAttempted",
                             "CreditsEarned", "EarnCreditIndicator", "SuccessInidicator", "VersionForGrade", "HasClassGrade", "ClassGrade", "GradeKey",
                             "GradeLetter", "GradePoints", "GradeDescription", "GradeSubgroup", "GradeGroup", "GradingBasisDescription")
```

#### Get list of course attributes in prereqs

```{r}
table(prerequisites$ValueFormalDescription)
```

#### Clean prereqs

```{r}
table(prerequisites$GradingBasisDescription, prerequisites$GradeLetter)
```

```{r}
# Clean data
p1 <- prerequisites %>%
  filter(EnrolledClassCount == 1,
         DropCount == 0,
         WithdrawCount == 0) %>%
  filter(!GradingBasisDescription %in% c("Administrative Grades", "Audit", "Complete Withdrawal", "Incomplete", "Non-Graded Component")) %>%
  # filter(HasClassGrade == 1) %>%     # Courses with "P" do not have ==1 here
  select(StudentID, TermSourceKey, PrimarySubject, PrimaryCatalogNumber, ClassSection, VersionForGrade, GradeLetter, GradePoints, GradeDescription, GradeSubgroup, GradeGroup, GradingBasisDescription, ValueFormalDescription)
```

#### Cast course attributes to dummy vars

```{r}
# Cast data
p2 <- reshape2::dcast(p1, 
                      StudentID + TermSourceKey + PrimarySubject + PrimaryCatalogNumber + ClassSection + VersionForGrade + GradeLetter + GradePoints + 
                        GradeDescription + GradeSubgroup + GradeGroup + GradingBasisDescription ~ ValueFormalDescription, 
                      value.var = "ValueFormalDescription")
```

```{r}
# Fix spaces in column names
names(p2) <- str_replace_all(names(p2), c(" " = ".",
                                          "," = ""))
```

```{r}
to_one <- function(x) {
  ifelse(x >= 1, 1, 0)
}

p3 <- p2 %>%
  mutate_at(.vars = vars(Afterwork.Program, Area.III.Core.course, Distance.Learning.Course, Dual.Enrollment, Exclude.from.Bronco.Budget,
                         Foundations.of.Natural.Physical.and.Applied.Sci, GEM.Common.Course.Number, Imaging.Sciences.BS.OPF.O, Military.Programs,
                         Mountain.Home.Program, Natural.Physical.and.Applied.Sciences, Online.Degree.Pathway.Lower.Div.Classes.OPF.O, Scientific.Ways.of.Knowing,
                         Weekend.University),
            .funs = to_one)
```

#### Ensure no duplicate or conflicting student-term-grades

```{r}
p3 %>%
  group_by(StudentID, TermSourceKey, PrimarySubject, PrimaryCatalogNumber) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
```


### Anonymize data

#### Pull hash table for anonymous ids

```{r}
# Convert prereq StudentID to a factor to be able to join
p4 <- p3 %>%
  mutate(StudentID = as.factor(StudentID))

p4 <- right_join(p4, student_anon_df, by = c("StudentID"))
```

### Remove PII

```{r}
p5 <- p4 %>%
  select(-StudentID)
```


```{r}
p5
```


### Prereq 1: BIOL 227

#### Choose only highest prereq grade and ensure no duplicates

```{r}
prereq_biol227_check <- p5 %>%
  filter(PrimarySubject == "BIOL" & PrimaryCatalogNumber == "227") %>%
  filter(!GradeSubgroup == "P") %>%
  group_by(StudentID_Anon) %>%
  slice(which.max(GradePoints)) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

head(prereq_biol227_check)
```

```{r}
prereq_biol227 <- p5 %>%
  filter(PrimarySubject == "BIOL" & PrimaryCatalogNumber == "227") %>%
  filter(!GradeSubgroup == "P") %>%
  group_by(StudentID_Anon) %>%
  slice(which.max(GradePoints))
```

#### Join with statics

```{r}
statics_w_bio227 <- left_join(statics, prereq_biol227, by = "StudentID_Anon")
```

```{r, fig.width = 9, fig.height = 7}

formula <- y ~ x


ggplot(statics_w_bio227, aes(y = NormalizedScore, x = GradePoints)) +
  geom_point(alpha = 0.75) +
  my_theme() +
  labs(title = "Relationship between Course and Prereq Performance", 
       subtitle = "Course: SU20 BIOL 228 \nPrereq: BIOL 227",
       x = "Grade Points in BIOL 227",
       y = "Grade in BIOL 228",
       caption = "Note: Excludes students without BIOL 227 GPA records \n Excludes P/F Grades, excludes AUD/W/CW/I") +
  geom_smooth(method = "lm") +
  stat_poly_eq(aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), 
               label.x.npc = "right", label.y.npc = 0.30,
               formula = formula, parse = TRUE, size = 5,
               color = "blue")

```





























