
# Setup

## Helper scripts

```{r}
source("./scripts/wrangling.R")
source("./scripts/visualizer.R")
source("./scripts/section_combiner.R")
```


## Set new parameters for reproducible pipeline

```{r}
# REQUIRED PARAMETERS
#course <- '1209_BIOL_227'
#course_directory <- paste0("/Users/bencroft/Google Drive/BIOL/", course, "/")

course <- 'BIOL_227_1209'
course_directory <- paste0("/Users/bencroft/Data/", course, "/")

# OPTIONAL PARAMETERS:

# Option 1: Combine two course sections
#course1_directory <- 
#course2_directory <- 
#course_combined_directory <- 
#course1_section_number <- 
#course2_section_number <- 
```

## Load libraries

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

library(ggplot2)
library(lubridate)
library(data.table)
library(scales)
library(stringr)
library(ggrepel)
library(ggridges)
library(ggpmisc)
library(kableExtra)
library(tidyr)
library(fastDummies)
library(dplyr)
```

## Load data

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
course_activity_input <- read.csv(paste0(course_directory, "course_activity.csv"), stringsAsFactors = FALSE)
course_item_activity_input <- read.csv(paste0(course_directory, "course_item_activity.csv"), stringsAsFactors = FALSE)
course_summary_input <- read.csv(paste0(course_directory, "course_summary.csv"), stringsAsFactors = FALSE)
forum_submissions_input <- read.csv(paste0(course_directory, "forum_submissions.csv"), stringsAsFactors = FALSE)
grade_center_input <- read.csv(paste0(course_directory, "grade_center.csv"), stringsAsFactors = FALSE)
student_course_summary_input <- read.csv(paste0(course_directory, "student_course_summary.csv"), stringsAsFactors = FALSE)
sis_input <- read.csv(paste0(course_directory, "sis.csv"), stringsAsFactors = FALSE)
#main_input <- read.csv(paste0(course_directory, "a4l_main.csv"))
#prereq_input <- readr::read_csv(paste0(course_directory, "prerequisites.csv"))
```


# Clean Data

## Change data types

```{r}
sis_input <- sis_input %>%
  mutate(StudentID = as.character(StudentID)) %>%
  mutate(AcademicPlan = as.character(AcademicPlan))
```

## Add grade order

```{r}
sis_input <- sis_input %>%
  mutate(GradeOrder = case_when(GradeLetter == "P" ~ 00,
                                 GradeLetter == "A+" ~ 01,
                                 GradeLetter == "A" ~ 02,
                                 GradeLetter == "A-" ~ 03,
                                 GradeLetter == "B+" ~ 04,
                                 GradeLetter == "B" ~ 05,
                                 GradeLetter == "B-" ~ 06,
                                 GradeLetter == "C+" ~ 07,
                                 GradeLetter == "C" ~ 08,
                                 GradeLetter == "C-" ~ 09,
                                 GradeLetter == "D+" ~ 10,
                                 GradeLetter == "D" ~ 11,
                                 GradeLetter == "D-" ~ 12,
                                 GradeLetter == "F" ~ 13,
                                 GradeLetter == "W" ~ 14,
                                 GradeLetter == "CW" ~ 15,
                                 GradeLetter == "I" ~ 16
                                 ))
```




## Get Student ID

```{r}
course_activity <- course_activity_input %>%
  mutate(StudentID = word(UserName, -1)) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\(", replacement = "")) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\)", replacement = ""))

course_item_activity <- course_item_activity_input %>%
  mutate(StudentID = word(UserName, -1)) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\(", replacement = "")) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\)", replacement = ""))

forum_submissions <- forum_submissions_input %>%
  mutate(StudentID = word(UserName, -1)) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\(", replacement = "")) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\)", replacement = ""))

grade_center <- grade_center_input %>%
  mutate(StudentID = word(Student, -1)) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\(", replacement = "")) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\)", replacement = ""))

student_course_summary <- student_course_summary_input %>%
  mutate(StudentID = word(Student, -1)) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\(", replacement = "")) %>%
  mutate(StudentID = gsub(StudentID, pattern = "\\)", replacement = ""))

sis <- sis_input
```




## Anonymize

```{r}
set.seed(1)

student_roster <- union(course_activity$StudentID, course_item_activity$StudentID)   
student_roster <- union(student_roster, forum_submissions$StudentID)    
student_roster <- union(student_roster, grade_center$StudentID)    
student_roster <- union(student_roster, student_course_summary$StudentID) 
student_roster <- union(student_roster, sis$StudentID)

# Create hash table for anonymous student ids
student_count <- length(unique(student_roster))
student_anon_ids <- sample(10000:99999, student_count, replace=FALSE)
student_anon_df <- data.frame(student_roster, student_anon_ids) %>%
  rename(StudentID = student_roster, StudentID_Anon = student_anon_ids) %>%
  mutate(StudentID = as.character(StudentID))

#sis <- sis %>% mutate(StudentID = as.character(StudentID))

course_activity <- left_join(course_activity, student_anon_df, by = c("StudentID"))
course_item_activity <- left_join(course_item_activity, student_anon_df, by = c("StudentID"))
forum_submissions <- left_join(forum_submissions, student_anon_df, by = c("StudentID"))
grade_center <- left_join(grade_center, student_anon_df, by = c("StudentID"))
student_course_summary <- left_join(student_course_summary, student_anon_df, by = c("StudentID"))
sis <- left_join(sis, student_anon_df, by = c("StudentID"))

# VALIDATE
sym_diff <- function(a,b) setdiff(union(a,b), intersect(a,b))

sis$StudentID_Anon %in% student_anon_df$StudentID_Anon
course_item_activity$StudentID_Anon %in% sis$StudentID_Anon
```


## Filter roles

```{r}
course_activity <- course_activity %>%
  filter(CourseRole == "Student")
course_item_activity <- course_item_activity %>%
  filter(CourseRole == "Student")
forum_submissions <- forum_submissions %>%
  filter(CourseRole == "Student")
```


## Remove PII

```{r}
course_activity <- course_activity %>%
  select(-StudentID, -UserName)
course_item_activityt <- course_item_activity %>%
  select(-StudentID, -UserName)
forum_submissions <- forum_submissions %>%
  select(-StudentID, -UserName)
grade_center <- grade_center %>%
  select(-StudentID, -Student)
student_course_summary <- student_course_summary %>%
  select(-StudentID)
sis <- sis %>%
  select(-StudentID, -EmployeeID)
```



## Relable First Term at Institution indicator

```{r}
sis <- sis %>%
  mutate(FirstTermAtInstitution = case_when(FirstTermAtInstitution == 1 ~ "First Term at BSU",
                                            FirstTermAtInstitution == 0 ~ "Not in First Term at BSU"))
```


## Rank AcademicLevel

```{r}
sis <- sis %>%
  mutate(AcademicLevel = case_when(AcademicLevel == "Freshman (10)" ~ "(10) Freshman",
                                    AcademicLevel == "Sophomore (20)" ~ "(20) Sophomore",
                                    AcademicLevel == "Junior (30)" ~ "(30) Junior",
                                    AcademicLevel == "Senior (40)" ~ "(40) Senior",
                                    AcademicLevel == "Post-Bacc Undergraduate (50)" ~ "(50) Post-Bacc Undergraduate",
                                    AcademicLevel == "Graduate (GR)" ~ "(60) Graduate"))
```



## Process Academic Plan data

### Remove Academic Subplan

```{r}
sis <- sis %>%
  select(-AcademicSubPlan) %>%
  distinct()
```

### One Hot Encode - Acad Plan

#### Get plans with more than 10 students

```{r paged.print=TRUE}
plan_counts <- sis %>%
  select(StudentID_Anon, AcademicPlan) %>%
  distinct() %>%
  group_by(StudentID_Anon, AcademicPlan) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count), AcademicPlan, StudentID_Anon) %>%
  ungroup() %>%
  group_by(AcademicPlan) %>%
  summarise(StudentsWithPlan = sum(Count)) %>%
  arrange(desc(StudentsWithPlan)) %>%
  ungroup()
plan_counts

plans_10_or_more <- plan_counts %>%
  filter(StudentsWithPlan >= 10) %>%
  mutate(AcademicPlan = as.character(AcademicPlan))
plans_10_or_more
```

#### Create new Academic Plan variable that wraps plans with n < 10 into a "low plan count" plan name

```{r}
sis2 <- sis %>%
  mutate(AcadPlan = ifelse(AcademicPlan %in% plans_10_or_more$AcademicPlan, AcademicPlan, "Plan w low count")) %>%
  select(AcadPlan, StudentID_Anon, everything()) %>%
  arrange(StudentID_Anon) 
```

#### Drop regular Academic Plan in favor of new plan (to spread the new variable, limiting new cols to plans with 10 or less)

```{r}
sis2 <- sis2 %>%
  select(-AcademicPlan)
```

#### Spread the new Academic Plan so each student is on 1 row

```{r}
sis2 %>%
  group_by(StudentID_Anon) %>%
  summarise(C = n()) %>%
  arrange(desc(C))
```

```{r}
sis2 %>%
  filter(StudentID_Anon == 61612)
```



#### Create plan dummies

```{r}
sis3 <- data.frame(cbind(sis2, dummies::dummy(sis2$AcadPlan))) # Append dummies for Academic Plan
```

#### Remove Plan column and collapse duplicates

```{r message=FALSE}
sis2_without_AcadPlanCol <- sis2 %>%
  select(-AcadPlan) 
grp_cols <- names(sis2_without_AcadPlanCol)

# Convert character vector to list of symbos
dots <- lapply(grp_cols, as.symbol)

sis4 <- sis3 %>%
  group_by_(.dots = dots) %>%
  summarise(across(starts_with("AcadPlan"), max)) %>%
  ungroup()
```

#### Test to see if there are student-term duplicates

```{r}
sis4 %>%
  group_by(TermSourceKey, ClassNumberUniqueDescription, StudentID_Anon) %>%
  summarise(C = n()) %>%
  arrange(desc(C))
```

## Jitter grades

```{r}
sis5 <- sis4 %>%
  mutate(ClassGrade_Jitter = jitter(as.numeric(ClassGrade))) %>%
  mutate(P1_ClassGrade_Jitter = jitter(as.numeric(P1_ClassGrade))) %>%
  mutate(P2_ClassGrade_Jitter = jitter(as.numeric(P2_ClassGrade))) %>%
  mutate(P3_ClassGrade_Jitter = jitter(as.numeric(P3_ClassGrade)))
```

#### Test to see if there are student-term duplicates

```{r}
sis5 %>%
  group_by(TermSourceKey, ClassNumberUniqueDescription, StudentID_Anon) %>%
  summarise(C = n()) %>%
  arrange(desc(C))
```

## Add BISQL to A4L data

```{r}
course_activity_sis <- left_join(course_activity, sis4, by = "StudentID_Anon")
course_item_activity_sis <- left_join(course_item_activity, sis4, by = "StudentID_Anon")
forum_submissions_sis <- left_join(forum_submissions, sis4, by = "StudentID_Anon")
grade_center_sis <- left_join(grade_center, sis4, by = "StudentID_Anon")
student_course_summary_sis <- left_join(student_course_summary, sis4, by = "StudentID_Anon")
```

```{r}
student_course_summary_sis %>%
  filter(!SISGradeLetter == "No Grade Recorded")
```


# Write Datasets

```{r}
write.csv(sis5, "/Users/bencroft/Google Drive/Projects/Course Analytics - BIOL 227 - Fall 2019/course_data_cleaned.csv", row.names = FALSE)
```

```{r}
min(sis5$ClassGrade_Jitter)
```


# Grade Analysis

```{r}
grade_center_sis$GradePercent = as.numeric(gsub("[\\%,]", "", grade_center_sis$GradePercent))
grade_center_sis %>%
  group_by(CourseItem)
```


```{r}
grade_center_sis %>%
  select(CourseItem, GradePercent)
```


```{r}
# sis4 %>%
  # select(starts_with("Plan10."))
```

## Data

```{r}
tmp <- sis4 %>%
  #dplyr::filter(Section %in% c(4001, 4002, 4201)) %>%
  filter(GradeSubgroup %in% c("A+/-", "B+/-", "C+/-", "D+/-", "F", "P")) %>%
  mutate(SuccessIndicator = as.factor(SuccessIndicator))

#tmp$AcademicLevel <- relevel(tmp$AcademicLevel, ref = "Sophomore (20)")
tmp$SuccessIndicator <- relevel(tmp$SuccessIndicator, ref = "No Success")
```


## Plans

### Plans Model 1: Linear Grade

```{r}
mod_PlansGrade <- lm(ClassGrade ~ AcadPlan..BIOL + AcadPlan..CJ.MIN + AcadPlan..FAMSTD.MIN + AcadPlan..HEALTHSCST + 
                       AcadPlan..HTHSCI.UND + AcadPlan..KINES.BS + AcadPlan..Plan.w.low.count + AcadPlan..PRE.BUS + AcadPlan..PRE.CJ +              AcadPlan..PRE.NURS + AcadPlan..PRE.RADSCI + AcadPlan..PSYCH.BS +AcadPlan..PSYCH.MIN + 
                       AcadPlan..RADSCI.BS + AcadPlan..SPANSH.MIN + AcadPlan..UNIV.UND,
           data = tmp)

summary(mod_PlansGrade)
```

### Plans THD

```{r}
#tukey_plans <- aov(mod_PlansGrade)
#tukey.test <- TukeyHSD(tukey_plans)
```


### Plans Model 2: Logistic Pass

```{r}
mod_PlansLikelihood <- glm(as.factor(SuccessIndicator) ~ 0 + AcadPlan..BIOL + AcadPlan..CJ.MIN + AcadPlan..FAMSTD.MIN + AcadPlan..HEALTHSCST + 
                       AcadPlan..HTHSCI.UND + AcadPlan..KINES.BS + AcadPlan..Plan.w.low.count + AcadPlan..PRE.BUS + AcadPlan..PRE.CJ +              AcadPlan..PRE.NURS + AcadPlan..PRE.RADSCI + AcadPlan..PSYCH.BS +AcadPlan..PSYCH.MIN + 
                       AcadPlan..RADSCI.BS + AcadPlan..SPANSH.MIN + AcadPlan..UNIV.UND,
           data = tmp, family = binomial)
summary(mod_PlansLikelihood)
```




## Prereq 

### Prereq Model 1: Linear Grade

```{r}
mod_PrereqGrade <- lm(ClassGrade ~ as.numeric(P1_ClassGrade) + as.numeric(P2_ClassGrade) + as.numeric(P3_ClassGrade),
           data = tmp)

summary(mod_PrereqGrade)
```

### Prereq Model 2: Logistic Pass

```{r}
mod_PrereqLikelihood <- glm(SuccessIndicator ~ as.numeric(P1_ClassGrade) + as.numeric(P2_ClassGrade) + as.numeric(P3_ClassGrade),
           data = tmp, family = binomial)
summary(mod_PrereqLikelihood)
```

## Academic Level

### Academic Level Model 1: Linear Grade

```{r}
mod_AcademicLevelGrade <- lm(ClassGrade ~ AcademicLevel,
           data = tmp)

summary(mod_AcademicLevelGrade)
```

### Academic Level Model 2: Logistic Pass

```{r}
mod_AcademicLevelLikelihood <- glm(SuccessIndicator ~ as.numeric(P1_ClassGrade) + as.numeric(P2_ClassGrade) + as.numeric(P3_ClassGrade),
           data = tmp, family = binomial)
summary(mod_AcademicLevelLikelihood)
```




























